Manuscript_Github_Pchabaudi_pir

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Changed total TPM of pir bar chart to a stacked chart with sub-families shown

@@

#Make and save z-scores
chab_z.list <- lapply(chab_tpm_avg_forZ.list,
                      tpm_to_zscore)

@@

```{r row_order_phaseograms}

#Getting the row order for each phaseogram for each experiment
# Want the data ordered by the highest zscore for each gene in each stage.
phaseogram_row_order <- lapply(chab_z.list, 
                               function(list_elements){
                                 zscore_df <- list_elements$zscores
                                 zscore_ordered <- melt(zscore_df, 
                                                        id = 'Geneid',
                                                        variable.name = 'stage',
                                                        value.name = 'zscore') %>% 
                                   group_by(Geneid) %>% 
                                   top_n(zscore, n = 1) %>% 
                                   arrange(stage, 
                                           desc(zscore)) %>%
                                   distinct(Geneid, 
                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
                                   filter(zscore > 0) #For some reason some remain with zero zscore
                                 # zscore_ordered <- zscore_ordered[order(
                                 #   match(
                                 #     zscore_ordered$stage, 
                                 #     order_vector
                                 #   )
                                 # ),]
                                 return(c(list_elements, 
                                          list(zscore_order = zscore_ordered)))
                               })

names(phaseogram_row_order) <- c('rMT', 'SBP')

phaseogram_time_order <- lapply(names(phaseogram_row_order), 
                                function(list_element){
                                  transmission_list <- getElement(phaseogram_row_order,
                                                                  list_element)
                                  order_by_time <- transmission_list$zscore_order %>% 
                                    mutate(stage = str_extract(stage,
                                                               pattern = '[[:digit:]]{2}h')) %>% 
                                    arrange(stage)
                                  return(list(zscore_order = order_by_time))
                                }
)
names(phaseogram_time_order) <- names(phaseogram_row_order)

@@

if(global_save_var){
 pdf(file = 'plots/phaseogram_pirgenes_rMTorder.pdf',
      height = 5,
      width = 5)
  draw_heatmap_zscore(zscore_list = chab_z.list,
                      gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                      filter_genes_vector = cir_id,
                      annotation_clade = 'SorL',
                      annotation_col = SorL_col,
                      threshold_tpm = 1,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = FALSE)
  dev.off()
}

if(global_save_var){
  pdf(file = 'plots/phaseogram_pirgenes_SBPorder.pdf',
      height = 5,
      width = 5)
  draw_heatmap_zscore(zscore_list = chab_z.list,
                      gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                      filter_genes_vector = cir_id,
                      annotation_clade = 'SorL',
                      annotation_col = SorL_col,
                      threshold_tpm = 1,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = TRUE)
  dev.off()
}

@@

```{r zscore_save}
chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
                        chab_z.list[[2]]$zscores, 
                        by = 'Geneid')
write.csv(chab_z.df,
          file = 'data/Pchabaudi_z.csv',
          row.names = FALSE)
write.table(x = phaseogram_time_order$rMT$zscore_order$Geneid,
            file = 'data/Pchabaudi_z_phaseOrder_rMT.list',
            row.names = FALSE, col.names = FALSE)
write.table(x = phaseogram_time_order$SBP$zscore_order$Geneid,
            file = 'data/Pchabaudi_z_phaseOrder_SBP.list',
            row.names = FALSE, col.names = FALSE)
```

>>>>

```{r all_pir_barchart}
cir_plot <- chab_tpm.df %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
         transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1],
         SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)]) %>% 
  group_by(Geneid, time, transmission, SorL) %>% 
  dplyr::summarise(tpm = median(tpm)) %>% 
  arrange(SorL, tpm)

y_scale <- seq(0, 
               signif(ceiling(
                 max(
                   group_by(cir_plot, time, transmission) %>%
                     dplyr::summarise(sum_tpm = sum(tpm)) %>%  .$sum_tpm
                   )
               ), 
               digits = 2), 
               length.out = 5)

(
  p <- ggplot(data = cir_plot, 
              aes(x = factor(time), 
                  y = tpm,
                  fill = factor(SorL))) +
    geom_col(colour = 'black',
             size = 0.1,
             position = 'stack') +
    scale_y_continuous(breaks = y_scale) +
    scale_fill_manual(values = SorL_col) +
    ylab(as.expression(bquote('Total '~italic(.('pir'))~'TPM'))) +
    xlab('Asexual blood stage development cycle') +
    theme_classic() +
    theme(axis.text.x=element_text(angle = 90, 
                                   vjust = 0.5,
                                   size = 7)) +
    facet_grid(~transmission, scales = 'free')
)

if(global_save_var){
  ggsave(filename = 'plots/allPir_BarChart.pdf',
          plot = p,
          dpi = 300,
          width = 20,
          height = 10,
          units = 'cm')
}

@@

#Make and save z-scores
# chab_z.list <- lapply(chab_tpm_avg_forZ.list,
#                       tpm_to_zscore)

@@

```{r row_order_phaseograms}
# 
# #Getting the row order for each phaseogram for each experiment
# # Want the data ordered by the highest zscore for each gene in each stage.
# phaseogram_row_order <- lapply(chab_z.list, 
#                                function(list_elements){
#                                  zscore_df <- list_elements$zscores
#                                  zscore_ordered <- melt(zscore_df, 
#                                                         id = 'Geneid',
#                                                         variable.name = 'stage',
#                                                         value.name = 'zscore') %>% 
#                                    group_by(Geneid) %>% 
#                                    top_n(zscore, n = 1) %>% 
#                                    arrange(stage, 
#                                            desc(zscore)) %>%
#                                    distinct(Geneid, 
#                                             .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
#                                    filter(zscore > 0) #For some reason some remain with zero zscore
#                                  # zscore_ordered <- zscore_ordered[order(
#                                  #   match(
#                                  #     zscore_ordered$stage, 
#                                  #     order_vector
#                                  #   )
#                                  # ),]
#                                  return(c(list_elements, 
#                                           list(zscore_order = zscore_ordered)))
#                                })
# 
# names(phaseogram_row_order) <- c('rMT', 'SBP')
# 
# phaseogram_time_order <- lapply(names(phaseogram_row_order), 
#                                 function(list_element){
#                                   transmission_list <- getElement(phaseogram_row_order,
#                                                                   list_element)
#                                   order_by_time <- transmission_list$zscore_order %>% 
#                                     mutate(stage = str_extract(stage,
#                                                                pattern = '[[:digit:]]{2}h')) %>% 
#                                     arrange(stage)
#                                   return(list(zscore_order = order_by_time))
#                                 }
# )
# names(phaseogram_time_order) <- names(phaseogram_row_order)
```

```{r read_z}
#Can read in the z scores to avoid re-doing all the above.
chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
#Make list object of the TPM values and the z-scores value by concatenating the 
# lists of the zscore dataframes and the TPM lists from above.
chab_z.list <- list('RMT' = c(chab_tpm_avg.list[['RMT']],
                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                           pattern = 'Geneid|rMT')])),
                    'SBP' = c(chab_tpm_avg.list[['SBP']],
                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                           pattern = 'Geneid|SBP')])))
phaseogram_time_order <- list(rMT = list(zscore_order = list()),
                              SBP = list(zscore_order = list()))

phaseogram_time_order$rMT$zscore_order$Geneid <- read.table(file = 'data/Pchabaudi_z_phaseOrder_rMT.list')[,1]
phaseogram_time_order$SBP$zscore_order$Geneid <- read.table(file = 'data/Pchabaudi_z_phaseOrder_SBP.list')[,1]

@@

if(global_save_var){
 pdf(file = 'plots/phaseogram_pirgenes_rMTorder.pdf',
      height = 5,
      width = 6)
  draw_heatmap_zscore(zscore_list = chab_z.list,
                      gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                      filter_genes_vector = cir_id,
                      annotation_clade = 'SorL',
                      annotation_col = SorL_col,
                      threshold_tpm = 1,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = TRUE)
  dev.off()
}

if(global_save_var){
  pdf(file = 'plots/phaseogram_pirgenes_SBPorder.pdf',
      height = 5,
      width = 6)
  draw_heatmap_zscore(zscore_list = chab_z.list,
                      gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                      # gene_ordering_vector = arrange(cir_clade_info, SorL)$Geneid,
                      filter_genes_vector = cir_id,
                      annotation_clade = 'SorL',
                      annotation_col = SorL_col,
                      threshold_tpm = 0,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = TRUE)
  dev.off()
}

@@

```{r zscore_save}
# chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
#                         chab_z.list[[2]]$zscores, 
#                         by = 'Geneid')
# write.csv(chab_z.df,
#           file = 'data/Pchabaudi_z.csv',
#           row.names = FALSE)
# write.table(x = phaseogram_time_order$rMT$zscore_order$Geneid,
#             file = 'data/Pchabaudi_z_phaseOrder_rMT.list',
#             row.names = FALSE, col.names = FALSE)
# write.table(x = phaseogram_time_order$SBP$zscore_order$Geneid,
#             file = 'data/Pchabaudi_z_phaseOrder_SBP.list',
#             row.names = FALSE, col.names = FALSE)
```
